Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-29 background.tex
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

We employ both tabular and pictorial means to 
articulate features of the proposed approaches and outline 
both strengths and weaknesses.
To ascertain the effectiveness of each proposal, we predominantly use 
three criteria that capture the behavior of the algorithms 
discussed:
\emph{efficiency}, \emph{overhead} and \emph{scalability}.
For each of these criteria, we use a coarse-grained,
three-level, gradation to denote how well the algorithms do,
namely \emph{low}, \emph{medium} and \emph{high}.
Each criterion has its own
interpretation of the above literal values. 
\emph{Efficiency} means how well the
algorithm performs its search and topology adaptation functions, without
affecting other characteristics of the networks 
(e.g. BitTorrent's optimal download performance). 
%%AD I am not sure what the above line says.. or connects. Something is 
%%	missing...
Efficiency is negatively effected every time an algorithm 
reduces its search scope.
For the \emph{overhead} criterion, we track the cost 
to compute and communicate peer join, message forwarding, 
and topology maintenance. 
For example, when exchanging route tables only amongst neighbors, 
we consider the traffic generated to be low.
%%AD Again I am not sure how the example above ties with the definition...
For the \emph{scalability} criterion, we consider system behavior 
with respect to the number of participating peers. 
We consider systems with low, medium, and high scalability 
to mean they can support hundreds, tenths of thousands, 
and millions of nodes, respectively.

>>>>

We employ both tabular and pictorial means to 
articulate features of the proposed approaches and outline 
both strengths and weaknesses.
To ascertain the effectiveness of each proposal, we predominantly use 
three criteria that capture the behavior of the algorithms 
discussed:
\emph{efficiency}, \emph{overhead} and \emph{scalability}.
For each of these criteria, we use a coarse-grained,
three-level, gradation to denote how well the algorithms do,
namely \emph{low}, \emph{medium} and \emph{high}.
Each criterion has its own
interpretation of the above literal values. 
\emph{Efficiency} means how well the
algorithm performs its search and topology adaptation functions, without
affecting other characteristics of the networks 
(e.g., BitTorrent's optimal download performance). 
%%AD I am not sure what the above line says.. or connects. Something is 
%%	missing...
Efficiency is negatively effected every time an algorithm 
reduces its search scope.
For the \emph{overhead} criterion, we track the cost 
to compute and communicate peer join, message forwarding, 
and topology maintenance. 
For example, when exchanging route tables only amongst neighbors, 
we consider the traffic generated to be low.
%%AD Again I am not sure how the example above ties with the definition...
For the \emph{scalability} criterion, we consider system behavior 
with respect to the number of participating peers. 
We consider systems with low, medium, and high scalability 
to mean they can support hundreds, tenths of thousands, 
and millions of nodes, respectively.

### 2014-01-29 conclusion.tex
2013.01.24 (Vassilis) Addressing Alex's comments in previous iteration for background and conclusion sections.
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

Research and development on application systems that follow the 
\p\ architectural paradigm, have flourished over the last decade.
By and large, such systems exploit the loose-coupling and self-organization 
of participating nodes to shape substrates upon which diverse applications 
can run; the latter offer services including, but not limited to, 
distributed naming, 
Internet telephony and video conferencing, 
storage and indexing, 
content and file sharing, 
web crawling and caching, 
event notification and subscribing/publishing.
These services are provided atop the 
best-effort infrastructure of today's Internet;
services have to comply with an array of key requirements and 
offer quality attributes that include node availability, robustness,
scalability, load-sharing, quality-of-service and user anonymity. 
%%AD rephrased below as above - too long sentence it was!
% and they have to effectively deal with an array of key quality attributes that
% they may need or have to offer node availability, robustness,
% scalability, load-sharing, quality-of-service and user anonymity.
%%
The difficulty to handle such characteristics effectively, 
emanates from the fact that
peers  operate at the edge of the Internet and no 
adequate attention is paid to the structure of the physical network 
during application-level network design. 
The deviation of the logical distributed overlay formation from the 
physical network yields suboptimal use of the underlying 
infrastructure and is known as the topology mismatch
problem.
%%AD rephrased what is below as above..
% The deviation of the structure of the overlay network from the
% optimal, according to the underlying physical structure, constitutes what is
% termed topology mismatch and the problem is known as the topology mismatch
% problem.

Development of \p--systems and applications
typically involve juggling a set of tradeoffs (e.g.,
choosing higher accuracy at the cost of increased system overhead).
% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
% choosing higher accuracy at the cost of increased system overhead.)  
In surveying more than
a decade's worth of research efforts aimed at solving the topology mismatch
problem in both unstructured and structured \p\ networks, we find this
tussle-of-tradeoffs property to hold.  
With regards to the three criteria we used --efficiency,
overhead, and scalability-- we find that none of the proposed solutions 
is superior to the others on all three fronts.  This is not surprising.
Nonetheless, by
\begin{inparaenum}[\itshape i\upshape)]
  \item presenting an analysis of each of the solutions,
including what distinguishes it from the
others and its pros and cons and
  \item offering a pictorial
comparison of how each solution fares in regards to others as far as
efficiency, overhead, and scalability are concerned,
\end{inparaenum}
we hope to have provided \p\
developers and researchers with enough insight and perspective;
as they face specific problems, they will be readily able to 
draw the best possible design decisions for their own application
environments.
%%AD sentence was too long - this "giant" thing to much ass-kissing... no.
% when facing a specific problem, they would 
% when they face a specific problem, th 
% so that depending
% on the particular problem in hand, they are able to make a
% correct choice amongst these surveyed here or, by firmly standing on
% the shoulders of giants, propose a new solution of their own.

>>>>

Research and development in application systems that follow the 
\p\ architectural paradigm have flourished for more than a decade now.
By and large, such systems exploit the loose-coupling and self-organization 
of participating nodes to shape substrates upon which diverse applications 
can run; the latter offer services including, but not limited to, 
distributed naming, 
Internet telephony and video conferencing, 
storage and indexing, 
content and file sharing, 
web crawling and caching, 
event notification and subscribing/publishing.
%%
These services are provided atop the 
best-effort infrastructure of today's Internet 
and they have to effectively deal with an array of key quality attributes that
they may have to offer, such as node availability, robustness,
scalability, load-sharing, quality-of-service and user anonymity.
%%
The difficulty to handle such features effectively, emanates from the fact that
peer nodes operate at the edge of the Internet and no adequate attention 
is paid to the structure of the physical network during the 
network formation at the application-level.
The deviation of the distributed logically-formed overlay from the 
physical network yields suboptimal use of the underlying 
infrastructure and is known as the topology mismatch
problem.

Development of \p--systems and applications
typically involve juggling a set of tradeoffs (e.g.,
choosing higher accuracy at the cost of increased system overhead).
% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
% choosing higher accuracy at the cost of increased system overhead.)  
In surveying more than
a decade's worth of research efforts aimed at solving the topology mismatch
problem in both unstructured and structured \p\ networks, we find this
tussle-of-tradeoffs property to hold.  
With regards to the three criteria we used --efficiency,
overhead, and scalability-- we find that none of the proposed solutions 
is superior to the others on all three fronts.  This is not surprising.
Nonetheless, by
\begin{inparaenum}[\itshape i\upshape)]
  \item presenting an analysis of each of the solutions,
including what distinguishes it from 
others as well as its advantages and disadvantages and 
  \item offering a pictorial
comparison of how each solution fares in regards to others as far as
efficiency, overhead, and scalability are concerned,
\end{inparaenum}
we hope to have provided \p\
developers and researchers with enough insight and perspective;
as they face specific problems, they will be readily able to 
draw the best possible design decisions for their own application
environments.
%%AD let's avoids "giants" etc...

### 2014-01-29 mard-survey.tex
2013.03.29 (Alex) Section on unstructured algorithms, massively reworked (from formatting to phrasing). Minor updates to background and conclusion sections.

\begin{abstract}
\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.  %% no commas here
They are implemented via an overlay network abstraction, on top 
of the best-effort IP infrastructure of the Internet. 
\p\ systems offer a plethora of desirable features to 
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, to name just a few.
Unfortunately, inherent weaknesses of \p\ systems, prevented 
applications from leveraging these features.
One major weakness, identified early on, is the \emph{topology mismatch problem}
between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability and efficiency of
applications built atop.
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the
suggested solutions and discuss the novelty as well as 
the  advantages and disadvantages of each such proposal.
Finally, we provide an analysis as well as pictorial comparisons 
of the reviewed approaches, since we aim to offer a comprehensive
reference for developers, system architects and researchers in the field.

>>>>

\begin{abstract}
\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.  %% no commas here
They are implemented via an overlay network abstraction, on top 
of the best-effort IP infrastructure of the Internet. 
\p\ systems offer a plethora of desirable features to 
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, to name just a few.
Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
applications from leveraging these features.
One major weakness, identified early on, is the \emph{topology mismatch problem}
between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability and efficiency of
applications built atop.
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the suggested solutions
by discussing their novelty, advantages and weaknesses.
Finally, we offer an analysis as well as pictorial comparisons 
of the reviewed approaches, since we aim to offer a comprehensive
reference for developers, system architects and researchers in the field.
